---
title: "Data Challenge"
author: "Zongyan Wang"
date: "March 15, 2016"
output: pdf_document
---
# How to use
To use the .Rmd file, it will send the user a request during the compiling to choose the data loaded. For the first request, please upload "LC\_biz\_all.csv". Please also place "importanceRank.txt" and "state\_latlon.csv" in your work directory. Thanks \
Please notice, this .Rmd file would also require clean\_f.R, plot\_f.R, analysis\_f.R.

# Upload packages for data clean and analysis
```{r}
require(plyr)  #data clean
require(dplyr) #data clean
require(tidyr) #data clean
require(ggplot2)  # visualzation
require(XML)   # web scriping
require(testthat) #test model
require(kernlab) # kernel and SVM
require(reshape2) 
require(caret) # Cross - Validation k fold
require(maps) # heatmap: map_data
require(ggmap) # for heatmap
```
# Read file
```{r}
# Read Data and function files
LC <- read.csv(file.choose(), header = T) #Please read LC_biz_all.csv
source("clean_f.R") # functions for cleaning the data
source("plot_f.R") # functions for plot(heatmap)
source("analysis_f.R") # functions for plot and data analysis
```

# Data clean
```{r}
lc <- perToN(LC)  # transfer percentage to numeric
lc <- replaceBlank_all(lc) # replace the blank
lc <- dateToNum(lc) # transform earliest_cr_line from date to numeric

```

# Data Summary
## Data overview
```{r}
dim(lc)               #variable size
hasNA_all(lc)            # variables contains missing values
length(hasNA_all(lc))
n.factor_all(lc)       # factor levels in variables among the data
length(n.factor_all(lc))
loan <- lc %>% 
  ddply(~loan_status, function(x) dim(x)[1]) %>%
  mutate(loan_status = reorder(x = loan_status,X = V1))
  
loanplot <- ggplot(loan, aes(x = loan_status, y = V1, fill = loan_status)) +
  geom_point(aes(color = loan_status),size = 4) +
  labs(y = "Amount", title = "Loan Status Summary") 
print(loanplot)
ggsave(filename = "loanplot.png", plot = loanplot, path = ".",  
       width = 10, height = 6, dpi = 400)
```


## Delete uninteresting variables
```{r}
#exclude loan_status, id, last payment day, next payment day, zip_code, issue_d
lc <- lc %>% select(-id,  -emp_title,
                    -last_pymnt_d, -next_pymnt_d,
                    -zip_code,-issue_d, -last_credit_pull_d)
```

## Paid off ratio over United States
```{r}
lc_fullypaid <- lc_ratio(lc, "Fully_Paid")
plot_fullypaid <- lc_heatmap(lc_fullypaid, state, location,
                             "Paid off Ratio over States")
print(plot_fullypaid)
ggsave(filename = "plot_fullypaid.png", plot = plot_fullypaid, path = ".",  
       width = 6, height = 6, dpi = 400)
```
## paid off and current (Good loan) ratio over United Status
```{r}
lc_current <- lc_ratio(lc, "Current")
lc_goodLoan <- merge(lc_current, lc_fullypaid, by = "addr_state")
lc_goodLoan <- lc_goodLoan %>% mutate(ratio = ratio.x + ratio.y) %>%
  select(-c(ratio.x,ratio.y))
plot_goodLoan <- lc_heatmap(lc_goodLoan, state, location,
                            "Good Loan (Paid Off + Current) Ratio over States")
print(plot_goodLoan)
ggsave(filename = "plot_goodLoan.png", plot = plot_goodLoan, path = ".",  
       width = 6, height = 6, dpi = 400)
```

# Data Analysis
## variable selection
```{r}
rank <- read.table("importanceRank.txt", header = T) # The list of important variables(importanceRank.txt) form GUIDE
rank <- rank %>% mutate(Variable = reorder(x = Variable, X = Scaled))
## plot rank
rankPlot <- rank %>% ggplot(aes(x = Variable, y = Scaled)) +
  geom_point(aes(color = Variable), size = 3)
print(rankPlot)
ggsave(filename = "rankPlot.png", plot = rankPlot, path = ".",  
       width = 10, height = 6, dpi = 400)
## Select the important data in original data
lc_i <- lc[,names(lc)%in%as.character(rank$Variable)]
lc_i$loan_status <- lc$loan_status
## Summary of new data
hasNA_all(lc_i)  # Variable has missing data
n.factor_all(lc_i)  # Levels of Variable which are categorical data
```

## logistic regression
```{r}
# resample the data with delinquency events
lc.sample <- sample_n(lc_i %>% subset(!loan_status %in% c("Current", "Fully_Paid")), 4000, replace = T)
lc_ii <- rbind(lc_i, lc.sample)

# weight other 10 times higher than good loan
lc_iii <- lc_i
lc_iii$weight <- 1
lc_iii$weight[!lc_i$loan_status %in% c("Current", "Fully_Paid")] <- 10

# Remove all the rows with NA
lc_i <- lc_i[-which(is.na(lc$percent_bc_gt_75)),]
# Data analysis
status <- as.character(levels(lc_i$loan_status))
# Set "Charged Off" to 1
log_charge <- log.loan(lc_i, status[1])
#model(log_charge, status[1])
# Set "Current" to 1
log_current <- log.loan(lc_i, status[2])
# Set "Default" to 1
log_default <- log.loan(lc_i, status[3])
# Set "Fully Paid" to 1
log_fd <- log.loan(lc_i, status[4])
# Set "In_Grace_Period" to 1
log_grace <- log.loan(lc_i, status[5])
# Set "Late_(16-30 days)" to 1
log_late1 <- log.loan(lc_i, status[6])
# Set "Late_(31 - 120 days)" to 1
log_late2 <- log.loan(lc_i, status[7])

# Plot the significant value
l.list <- list(log_charge, log_current, log_default, log_fd, log_grace, log_late1, log_late2)
plot_coeff <- coeff_plot(l.list,status, sig = F)
plot_significant <- coeff_plot(l.list, status, sig = T)
print(plot_coeff)
ggsave(filename = "plot_coeff.png", plot = plot_coeff, path = ".",  
       width = 10, height = 6, dpi = 400)
print(plot_significant)
ggsave(filename = "plot_significant.png", plot = plot_significant, path = ".",  
       width = 10, height = 6, dpi = 400)
```
## multi-level Logistic regression model accuracy(without weight)
```{r, warning= F}
## in-sample error
l.list <- glm_list(lc_i, status)
in_sample_pred <- predict_glm(l.list, lc_i, status)
in_sample_error <- 1 - sum(in_sample_pred == lc_i$loan_status)/length(lc_i[,1])
in_sample_error
# plot classification result
plot_mis1 <- mis_plot(true = lc_i$loan_status, pred = in_sample_pred)
print(plot_mis1)
ggsave(filename = "plot_mis1.png", plot = plot_mis1, path = ".",  
       width = 10, height = 6, dpi = 400)
```
## multi-level Logistic regression model accuracy(resample)
```{r}
## in-sample error
newdata <- lc_i
l.list <- glm_list(lc_ii, status)
in_sample_pred <- predict_glm(l.list, newdata, status)
in_sample_error <- 1 - sum(in_sample_pred == lc_i$loan_status)/length(lc_i[,1])
in_sample_error
# plot classification result
plot_mis2 <- mis_plot(true = lc_i$loan_status, pred = in_sample_pred)
print(plot_mis2)
ggsave(filename = "plot_mis2.png", plot = plot_mis2, path = ".",  
       width = 10, height = 6, dpi = 400)

```


## multi-level Logistic regression model accuracy(weight)
```{r}
## in-sample error
newdata <- lc_i
newdata$weight <- 1
l.list <- glm_list(lc_iii, status)
in_sample_pred <- predict_glm(l.list, newdata, status)
in_sample_error <- 1 - sum(in_sample_pred == lc_iii$loan_status)/length(lc_iii[,1])
in_sample_error
# plot classification result
plot_mis3 <- mis_plot(true = lc_i$loan_status, pred = in_sample_pred)
print(plot_mis3)
ggsave(filename = "plot_mis3.png", plot = plot_mis3, path = ".",  
       width = 10, height = 6, dpi = 400)


```
## Kernel methods
```{r}
lc_svm_i <- prep_svm(lc_i)  
# resample data for svm
lc_svm_ii <- prep_svm(lc_ii)
# sample SVM model"
svmmodel <- svm(lc_svm_i,"Fully_Paid")

  #our SVM model
newdata <- as.matrix(lc_svm_i%>% select(-loan_status))
kern.list <- kern_list(lc_svm_i)
pred_svm <- pred.svm(kern.list, newdata, status = status)

in_sample_error_kern <- 1 - sum(pred_svm == lc_i$loan_status)/length(lc_i[,1])
in_sample_error_kern
# plot classification result
plot_mis_kern <- mis_plot(true = lc_i$loan_status, pred = pred_svm)
print(plot_mis_kern)
ggsave(filename = "plot_mis_kern.png", plot = plot_mis_kern, path = ".",  
       width = 10, height = 6, dpi = 400)


```
# SVM for data after resampling
```{r}
#our SVM model
newdata <- as.matrix(lc_svm_i%>% select(-loan_status))
kern.list2 <- kern_list(lc_svm_ii)
pred_svm2 <- pred.svm(kern.list2, newdata, status = status)

in_sample_error_kern2 <- 1 - sum(pred_svm == lc_i$loan_status)/length(lc_i[,1])
in_sample_error_kern2
# plot classification result
plot_mis_kern2 <- mis_plot(true = lc_i$loan_status, pred = pred_svm2)
print(plot_mis_kern2)
ggsave(filename = "plot_mis_kern2.png", plot = plot_mis_kern2, path = ".",  
       width = 10, height = 6, dpi = 400)
```